High-Throughput Computing on \\ High-Performance Platforms: A Case Study

### 2017-09-10 PandaEvolution.tex
ongoing cleanup.

\ifreview
Comments to address in this section:
\begin{enumerate}
	\color{red} 
	\item REVIEWER 2: the technical discussion of TITAN and WLCG differences
	is abruptly stopped by the description of Monte Carlo event generation
	and detector simulation (3A).\mtnote{I read the section twice and I did
	not see such an abrupt transition.}
	\item REVIEWER 2: The single listing (3B) of similar projects within the
	collaboration does not detail whether this is derived or separate
	work.\mtnote{Added a sentence to clarify that PanDA Broker operates
	within the existing ATLAS software infrastructure. Specified that the
	ATLAS team has ported PanDA Broker to other supercomputers.}
	\item REVIEWER 2: How does the TITAN environment prevent usage of DTNs as
	thin proxies? Is connectivity possible from compute node to DTNs via MPI,
	message queues, sockets or other means? Are the restrictions a
	shortcomming of TITAN, or the majority of HPC environments? Given the
	proclaimed importance of HPC resources, a generic evaluation seems
	needed.\mtnote{Added a sentence explaining that DTNs are not part of
	Titan's compute nodes.}
	\item REVIEWER 2: Why was the PanDA Broker designed orthogonal to the
	pilot model, performing both resource acquisition and workload
	execution?\mtnote{We explain that PanDA Broker is a prototype and now we
	are developing NGE.} How do pilots conflict with backfill slots, when the
	former are designed to acquire resources independent of workload
	requirements?\mtnote{AFAIK, we never state that pilot conflict with
	backfill. We argue that the latter mitigates the absence of the former.}
	Why does each Broker instance query and use backfill separately? Given
	that this seems to be a major bottleneck, the choice of such a design
	begs for motivation.\mtnote{Extended a paragraph to explain that PanDA
	Broker supported rapid deployment and incremental optimization while
	incurring in functional limitations.}
	\item REVIEWER 2: Some facts about WLCG seem outdated:
	\begin{itemize}
		\item The number of 30,000 unused TITAN cores (§IIIB) is equated to
		roughly 10\% of the WLCG. However, WLCG currently has 591,282 cores
		[REBUSCPU].
		https://wlcg-rebus.cern.ch/apps/capacities/federations/\mtnote{Done.}
		\item WLCG sites are claimed to provide only 8 cores per worker node
		(§IIIB). WLCG has a significant number of sites providing 16 or more
		cores per worker node [REBUSCPU].
		https://wlcg-rebus.cern.ch/apps/capacities/federations/\mtnote{This
		would require fairly deep rewriting on Section IVB. I edited the
		sentence referring to 8 cores eliminating the reference to compute
		nodes. Less than ideal but I did not find a better solution. Any
		suggestion?}
	\end{itemize}
	\item REVIEWER 2: Schematic figures (Fig 1, Fig 2) are overly complex,
	while their captions are not sufficient to understand them. The numbering
	is only partially explained in the text.\mtnote{Added a reference to
	where the numbering of the figures are explained in the text. I
	double-checked and the numbering of both figure is fully explained in the
	text.}
\end{enumerate}
\fi

>>>>

% \ifreview
% Comments to address in this section:
% \begin{enumerate}
% 	\color{red} 
% 	\item REVIEWER 2: the technical discussion of TITAN and WLCG differences
% 	is abruptly stopped by the description of Monte Carlo event generation
% 	and detector simulation (3A).\mtnote{I read the section twice and I did
% 	not see such an abrupt transition.}
% 	\item REVIEWER 2: The single listing (3B) of similar projects within the
% 	collaboration does not detail whether this is derived or separate
% 	work.\mtnote{Added a sentence to clarify that PanDA Broker operates
% 	within the existing ATLAS software infrastructure. Specified that the
% 	ATLAS team has ported PanDA Broker to other supercomputers.}
% 	\item REVIEWER 2: How does the TITAN environment prevent usage of DTNs as
% 	thin proxies? Is connectivity possible from compute node to DTNs via MPI,
% 	message queues, sockets or other means? Are the restrictions a
% 	shortcomming of TITAN, or the majority of HPC environments? Given the
% 	proclaimed importance of HPC resources, a generic evaluation seems
% 	needed.\mtnote{Added a sentence explaining that DTNs are not part of
% 	Titan's compute nodes.}
% 	\item REVIEWER 2: Why was the PanDA Broker designed orthogonal to the
% 	pilot model, performing both resource acquisition and workload
% 	execution?\mtnote{We explain that PanDA Broker is a prototype and now we
% 	are developing NGE.} How do pilots conflict with backfill slots, when the
% 	former are designed to acquire resources independent of workload
% 	requirements?\mtnote{AFAIK, we never state that pilot conflict with
% 	backfill. We argue that the latter mitigates the absence of the former.}
% 	Why does each Broker instance query and use backfill separately? Given
% 	that this seems to be a major bottleneck, the choice of such a design
% 	begs for motivation.\mtnote{Extended a paragraph to explain that PanDA
% 	Broker supported rapid deployment and incremental optimization while
% 	incurring in functional limitations.}
% 	\item REVIEWER 2: Some facts about WLCG seem outdated:
% 	\begin{itemize}
% 		\item The number of 30,000 unused TITAN cores (§IIIB) is equated to
% 		roughly 10\% of the WLCG. However, WLCG currently has 591,282 cores
% 		[REBUSCPU].
% 		https://wlcg-rebus.cern.ch/apps/capacities/federations/\mtnote{Done.}
% 		\item WLCG sites are claimed to provide only 8 cores per worker node
% 		(§IIIB). WLCG has a significant number of sites providing 16 or more
% 		cores per worker node [REBUSCPU].
% 		https://wlcg-rebus.cern.ch/apps/capacities/federations/\mtnote{This
% 		would require fairly deep rewriting on Section IVB. I edited the
% 		sentence referring to 8 cores eliminating the reference to compute
% 		nodes. Less than ideal but I did not find a better solution. Any
% 		suggestion?}
% 	\end{itemize}
% 	\item REVIEWER 2: Schematic figures (Fig 1, Fig 2) are overly complex,
% 	while their captions are not sufficient to understand them. The numbering
% 	is only partially explained in the text.\mtnote{Added a reference to
% 	where the numbering of the figures are explained in the text. I
% 	double-checked and the numbering of both figure is fully explained in the
% 	text.}
% \end{enumerate}
% \fi

### 2017-09-10 overall.tex
ongoing cleanup.

    \item REVIEWER 2: Why does Fig 3 show an increase of absolute and
    relative efficiency in Aug-16, when the text claims this in Sep-16 due to
    more PanDA Brokers?\mtnote{Fixed.}

    \item REVIEWER 2: Why are only 20 PanDA Brokers deployed if this is a
    limiting factor? Why is it questionable that more PanDA Brokers would
    increase efficiency? If only 5/6 of resources below the mean are
    unusable, this implies an efficiency of almost 60\% being
    reachable.\mtnote{From our text: ``further research is ongoing to
    understand whether an even greater number of brokers would yield even
    greater efficiency. This research focuses on evaluating the overheads of
    input/output files staging, including its impact on DTNs, and on an
    alternative design of PanDA Broker that enables the concurrent submission
    of multiple MPI scripts''. I believe this is an answer good enough.}

    \item REVIEWER 2: Given an expected increase in resource need of 10-100,
    how is the capacity of 4\% WLCG significant? How does this relate to the
    manpower required to develop and maintain a custom solution per HPC site?
    How many TITAN resources are actually available to HEP, given the use by
    other domains? The magnitude of capacities presented as available is in
    major contrast to the requirements sketched in the introduction.

    \item REVIEWER 2: How does the poor metadata read performance of
    1500-6300s (25\%-100\% of minimum walltime!) affect efficiency? Is it
    factored into the choice of walltime requests?\mtnote{Thanks to Sergey's
    input I corrected the figure previously reported.} Is it applicable only
    to TITAN, to HPC in general, or even to regular WLCG sites?\mtnote{This
    is clearly out of scope for this paper.} Are there solutions at
    application, e.g. the use of packaged python scripts and
    libraries?\mtnote{Added the following: ``While these figures are specific
    to OLCF and TItan, reengineering on AthenaMP could further improve
    startup time on every platform.''} The information presented is
    insufficient to estimate whether this issue applies to other environments
    and usecases as well, or is simply due to suboptimal
    implementation/configuration.

    \item REVIEWER 3: The HEP-SPEC06 benchmark, which is used as a
    ``computing power'' measure in LCG is mostly integer-artihmetic based: FP
    operations accounts for \~10\% of the average workload (see [1], section
    ``Why do you use `all\_cpp' instead of SPECint or SPECfp?'') While this
    figure certainly comes from averaging over many different workloads
    across all experiments, it seems a bit surprising that FP scheduling
    contention is responsible for 50\% decrease in compute efficiency on
    Titan when compared to other WLCG sites (paper page 6).\mtnote{Thanks to
    Sergey's input I downplayed the role played by FP and up-played the role
    played by new CPUs.}
\end{enumerate}
\fi

>>>>

    % \item REVIEWER 2: Why does Fig 3 show an increase of absolute and
    % relative efficiency in Aug-16, when the text claims this in Sep-16 due to
    % more PanDA Brokers?\mtnote{Fixed.}

    % \item REVIEWER 2: Why are only 20 PanDA Brokers deployed if this is a
    % limiting factor? Why is it questionable that more PanDA Brokers would
    % increase efficiency? If only 5/6 of resources below the mean are
    % unusable, this implies an efficiency of almost 60\% being
    % reachable.\mtnote{From our text: ``further research is ongoing to
    % understand whether an even greater number of brokers would yield even
    % greater efficiency. This research focuses on evaluating the overheads of
    % input/output files staging, including its impact on DTNs, and on an
    % alternative design of PanDA Broker that enables the concurrent submission
    % of multiple MPI scripts''. I believe this is an answer good enough.}

    \item REVIEWER 2: Given an expected increase in resource need of 10-100,
    how is the capacity of 4\% WLCG significant? How does this relate to the
    manpower required to develop and maintain a custom solution per HPC site?
    How many TITAN resources are actually available to HEP, given the use by
    other domains? The magnitude of capacities presented as available is in
    major contrast to the requirements sketched in the introduction.\jhanote{MT to address this via a few sentences in Section IV.}

    % \item REVIEWER 2: How does the poor metadata read performance of
    % 1500-6300s (25\%-100\% of minimum walltime!) affect efficiency? Is it
    % factored into the choice of walltime requests?\mtnote{Thanks to Sergey's
    % input I corrected the figure previously reported.} Is it applicable only
    % to TITAN, to HPC in general, or even to regular WLCG sites?\mtnote{This
    % is clearly out of scope for this paper.} Are there solutions at
    % application, e.g. the use of packaged python scripts and
    % libraries?\mtnote{Added the following: ``While these figures are specific
    % to OLCF and TItan, reengineering on AthenaMP could further improve
    % startup time on every platform.''} The information presented is
    % insufficient to estimate whether this issue applies to other environments
    % and usecases as well, or is simply due to suboptimal
    % implementation/configuration.

    % \item REVIEWER 3: The HEP-SPEC06 benchmark, which is used as a
    % ``computing power'' measure in LCG is mostly integer-artihmetic based: FP
    % operations accounts for \~10\% of the average workload (see [1], section
    % ``Why do you use `all\_cpp' instead of SPECint or SPECfp?'') While this
    % figure certainly comes from averaging over many different workloads
    % across all experiments, it seems a bit surprising that FP scheduling
    % contention is responsible for 50\% decrease in compute efficiency on
    % Titan when compared to other WLCG sites (paper page 6).\mtnote{Thanks to
    % Sergey's input I downplayed the role played by FP and up-played the role
    % played by new CPUs.}
\end{enumerate}
\fi
